# Reproducible Research: Peer Assessment1

## Loading and preprocessing the data
Read the data from the file *"activity.csv"*
```{r, echo=TRUE}
Data<-read.csv("activity.csv")
```
Chage the laguage to english
```{r language,echo=TRUE}
 Sys.setlocale("LC_TIME","English")
````

Select the steps from the Data and transform as numeric value
```{r steps, echo=TRUE}
steps<-as.numeric(as.character(Data[,1]))
```
Select the date from Data and transform as Date value.
```{r date,echo=TRUE}
date<-as.Date(Data[,2],"%Y-%m-%d")
```
Select the interval from Data and transform as numeric value
```{r interval,echo=TRUE}
interval<-as.numeric(as.character(Data[,3]))
```




## What is mean total number of steps taken per day?

Create a frame with the values of steps and date.
```{r framestepsdate,echo=TRUE}
datastepsdate<-data.frame(steps,date)
```

Combine the datastepsdate frame per day
```{r stepsday,echo=TRUE}
library(plyr)
stepsday<-ddply(datastepsdate,.(date),summarize,stepsperday=sum(steps))
```

Plot the histogram of the total number of steps taken each day.
```{r histogram, echo=TRUE}
hist(stepsday$stepsperday, col="red",xlab="number of steps", main="Histogram of steps taken per day")
```

Calculate the **mean**  and **median** total number of the steps taken per day
```{r meanandmedian, echo=TRUE}
meanandmedian<-ddply(datastepsdate,.(date),summarize,mean=mean(steps,na.rm=TRUE),median=median(steps, na.rm=TRUE))
print(meanandmedian)
```

## What is the average daily activity pattern?

Create a frame with the values of steps and interval.
```{r stepsinterval,echo=TRUE}
stepsinterval<-data.frame(steps, interval)
```

Combine the stepsinterval frame per interval.
```{r stepsperinterval,echo=TRUE}
stepsperinterval<-ddply(stepsinterval,.(interval),summarize,stepsperinterval=mean(steps,na.rm=TRUE))
```

1. Plot the 5-minute interval and the average number of steps taken, averaged across all days.

```{r plotstepsperinterval, echo=TRUE}
plot(stepsperinterval, type="l", ylab="average of number of steps" )
```

Obtain the maximum number of steps.
```{r max,echo=TRUE}
maxsteps<-max(steps, na.rm=TRUE)
```
Find the location of the interval with the step maximum.
```{r findlocationmaxstep,echo=TRUE}
locationmaxstep<-stepsinterval[stepsinterval$steps %in% c(maxsteps),]
```

2. The 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps is:
```{r locationmaxstep,echo=TRUE}
locationmaxstep[[2]]
```


## Imputing missing values

1. The total number of missing values in the dataset.
```{r numNAs,echo=TRUE}
stepsinf<-summary(steps)
stepsinf[["NA's"]]
```

2. For filling in all of the missing values in the dataset. We are going to use the mean/median for that day, or the mean for that 5-minute interval.

```{r fillingdata,echo=TRUE}
Newsteps<-steps                                                 ## Create a new list of steps
for(i in 1:length(steps)){                                      ## The for loop run over the list of steps
        sp<-steps[i]                                            ## select a specific number of steps from the list
        if(is.na(sp)){                                          ## check if the speci number is NA
                day<-date[i]                                    ## select the date from that specific number
                MM<-meanandmedian[meanandmedian$date %in% c(day),] ## give the value of mean and median
                                                                   ## for the specific date
                Mean<-MM[["mean"]]                              ## select the mean value for the specific date
                Median<-MM[["median"]]                          ## select the median value for the specific date
                if(is.na(Mean)){                                ## check if the mean is NA
                        if(is.na(Median)){                      ## check if the median is NA
                                Int<-interval[i]                ## select the interval from that specific number
                                IS<-stepsperinterval[stepsperinterval$interval %in% c(Int),] ##give the value of
                                                                                             ## average of interval
                                stp<-IS[["stepsperinterval"]]   ## select the steps per interterval for that interval
                                }
                        if (!is.na(Median)){                    ## check if the median is not NA.
                                stp<-Median                     ## call stp the median value. 
                                }
                        Newsteps[i]<-stp                        ## assigne the value stp(stepsperinterval/median) to
                                                                ## new steps list
                        }
                if(!is.na(Mean)){                               ## check if the mean is not NA.
                        Newsteps[i]<-Mean                       ## assigne the value of mean to the new steps list
                        }
         }
        if (!is.na(sp)){                                        ## check if the specific step value is not NA.
                Newsteps[i]<-steps[i]                           ## assigne the value of specific to the new list
        }
}
```
3. New dataset that is equal to the original dataset but with the missing data filled in

```{r NewData,echo=TRUE}
Newdata<-data.frame(Newsteps,date,interval)
```

4. a. Histogram of the total number of steps taken each day
```{r histogramNewdata,echo=TRUE}
stepsdaynewdata<-ddply(Newdata,.(date),summarize,totalstepsperday=sum(Newsteps))
hist(stepsdaynewdata$totalstepsperday, col="red",,xlab="number of steps taken per day", main="Histogram of steps taken per day without NA's")
```
   
   b. Calculate the **mean** and **median** total number of steps taken per day. 
```{r Newdatameanandmedian, echo=TRUE}
Newdatameanmedian<-ddply(Newdata,.(date),summarize,mean=mean(Newsteps),median=median(Newsteps))
print(Newdatameanmedian)
```
   
   c. **Do these values differ from the estimates from the first part of the assignment?**
   
*This values only differ from the estimates from the first part in the information for the missing values, because the missing values are filled in with the corresponding information mean/median or average of 5-minutes interval. The rest of information remain equal.* 

   **What is the impact of imputing missing data on the estimates of the total daily number of steps?**
      
*The impact of imputing missing data on the estimates of the total daily number of steps does not change when compared to the original the estimate of the total daily number of steps.*
           


## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r weekdays,echo=TRUE}
days<-weekdays(date)                     ## change the date list to the names of weekday
weekdayandweekend<-days                  ## create a new list
for(i in 1:length(days)){                ## the for loop run over the days of the new list
        day<-days[i]                     ## select a specific day from the new list
        if(day=="Monday"| day=="Tuesday"| day=="Wednesday" |day=="Thursday" | day=="Friday"){ ##check if the day is
                                                                                              ##weekday
                weekdayandweekend[i]<-"weekday"  ## assign the tagline weekday to the day
        }
        if(day=="Saturday" | day=="Sunday"){ ## check if the day is weekend
                weekdayandweekend[i]<-"weekend" ## assign the tagline weekend to the day
        }
}
```
2. Create a frame with the values of steps and weekdays or weekend.
```{r framestepsdays,echo=TRUE}
DT<-data.frame(Newsteps,weekdayandweekend,interval)
```

Combine the DT frame per day
```{r DTstepsdays,echo=TRUE}
FinalData<-aggregate(Newsteps ~ weekdayandweekend +interval , data = DT, mean)
```

Plot a time series of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.
```{r,echo=TRUE}
library(lattice)
xyplot(Newsteps ~ interval |weekdayandweekend , data = FinalData, layout = c(1,2),type="l", ylab="Number of steps", xlab="Interval")
```